---
title: "Best Presentation"
subtitle: "Stats 115 Final Project"
author: "Charles Gao<br>Andrew Hwang<br>Jason Huffstutter<br>Andrew Hahm"
format: revealjs
editor: visual
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(bayesrules)
library(rstanarm)
library(bayesplot)
library(tidyverse)
library(broom.mixed)
library(tidybayes)

drugs <- read_csv("data/drugs.csv")
```

## Background & Motivation

TO DO

## Our Research Question

-   How do years of economic hardship (2008) affect alcohol rates across all age groups?
-   Did early state-level marijuana legalization in 2012 (Washington, Colorado) and 2014 (Alaska, Oregon) represent a causal driver of increased usage, or did usage rates simply follow the national trajectory?

## Dataset Overview

::: {style="font-size: 0.7em;"}
```{r}
cat("Dimensions:", nrow(drugs), "observations ×", ncol(drugs), "variables\n")
cat("Years:", min(drugs$Year), "-", max(drugs$Year), "\n")
cat("States:", length(unique(drugs$State)), "\n")
```

**Key Variables:**

-   **Demographics**: State, Year, Population by age group (12-17, 18-25, 26+)
-   **Substances**: Alcohol, Marijuana, Tobacco, Cocaine
-   **Metrics**: Totals (counts) and Rates (per 1000 population)
-   **Time frames**: Past month, past year usage

**Age Groups:**

-   12-17 (adolescents)
-   18-25 (young adults)
-   26+ (adults)
:::

## Marijuana Trends by Age

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Marijuana.Used Past Month.12-17`,
         `Rates.Marijuana.Used Past Month.18-25`,
         `Rates.Marijuana.Used Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Marijuana Use Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group shows an increasing rate per 1,000 over time and remains significantly higher than the other two age groups.
:::

## Alcohol Disorder Trends

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Alcohol.Use Disorder Past Year.12-17`,
         `Rates.Alcohol.Use Disorder Past Year.18-25`,
         `Rates.Alcohol.Use Disorder Past Year.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Alcohol Use Disorder Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group still shows the highest alcohol disorder rate per 1,000, but rates for all three age groups gradually decrease over time.
:::

## Tobacco Trends

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Tobacco.Cigarette Past Month.12-17`,
         `Rates.Tobacco.Cigarette Past Month.18-25`,
         `Rates.Tobacco.Cigarette Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Tobacco Use Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group shows the highest tobacco usage rate per 1,000. Rates for all three age groups gradually decline over time, but at different speeds, with the 18–25 group eventually converging with the 26+ group.
:::

## Top States: Marijuana Use

```{r fig.width=10, fig.height=7}
drugs %>%
  filter(Year == max(Year)) %>%
  arrange(desc(`Rates.Marijuana.Used Past Month.18-25`)) %>%
  slice(1:15) %>%
  ggplot(aes(x = reorder(State, `Rates.Marijuana.Used Past Month.18-25`), 
             y = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 15 States: Marijuana Use (18-25)", x = "", y = "Rate per 1000")
```

## Distribution: Marijuana Use

```{r fig.width=10, fig.height=6}
drugs %>%
  ggplot(aes(x = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_histogram(bins = 30, fill = "seagreen", alpha = 0.7) +
  geom_density(aes(y = after_stat(count)), size = 1.2, color = "darkgreen") +
  labs(title = "Distribution of Marijuana Use Rates (18-25)", x = "Rate per 1000")
```

## Correlation Matrix

```{r fig.width=12, fig.height=10}
cor_data <- drugs %>%
  select(Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Alcohol = `Rates.Alcohol.Use Disorder Past Year.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`)

cor_matrix <- cor(cor_data, use = "complete.obs")

cor_df <- as.data.frame(as.table(cor_matrix))
names(cor_df) <- c("Var1", "Var2", "Correlation")

ggplot(cor_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white", size = 1) +
  geom_text(aes(label = round(Correlation, 2)), size = 8) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Between Substances (18-25)") +
  theme_minimal(base_size = 16) +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.text.y = element_text(size = 14))
```

::: {style="font-size: 0.8em;"}
**Two distinct camps:** Cocaine & Marijuana users (r = 0.55) vs Alcohol & Tobacco users (r = 0.73)

**Key insight:** Marijuana and Tobacco negatively correlated (r = -0.27) - both are smoking activities, individuals choose one over the other
:::

## All Substance Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  select(Year, 
         Alcohol = `Rates.Alcohol.Use Past Month.18-25`,
         Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`) %>%
  pivot_longer(cols = -Year, names_to = "Substance", values_to = "Rate") %>%
  ggplot(aes(x = Year, y = Rate, color = Substance)) +
  geom_smooth(se = TRUE) +
  facet_wrap(~Substance, scales = "free_y") +
  labs(title = "All Substances (18-25 Age Group)") +
  theme(legend.position = "none")
```

## Question 1: Economic Hardship

**How do years of economic hardship (2008) affect alcohol rates across all age groups?**

Initially, we thought that many people would cope with hardship by abusing alcohol (as seen in popular culture), but we also figured that many households might also employ a tighter monetary policy to save some money.

## Question 2: Marijuana Legalization

**Did early state-level marijuana legalization in 2012 (Washington, Colorado) and 2014 (Alaska, Oregon) lead to higher marijuana-use rates among ages 18–25, beyond the broader national trend?**

Our exploratory analysis suggested that marijuana use among **18–25-year-olds** rose after legalization. We then asked whether legalization acted as an additional driver of growth in early-adopter states, or whether usage was simply increasing along with the national trajectory.

## Legalization Timeline

::: {style="font-size: 0.75em;"}
**Early Adopter States:**

-   **2012**: Washington & Colorado legalized recreational marijuana
-   **2014**: Alaska & Oregon legalized recreational marijuana

**Our Analysis:**

Compare marijuana usage rates in these 4 early-adopter states vs. all other states to identify:

1.  Pre-legalization baseline differences
2.  Post-legalization divergence (if any)
3.  Whether legalization accelerated existing trends
:::

## Marijuana Rates: Early Adopters vs Others

```{r fig.width=12, fig.height=7}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

drugs %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Year, Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = Year, y = Avg_Rate, color = Group, group = Group)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_vline(xintercept = 2012, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "orange", size = 1) +
  annotate("text", x = 2012, y = Inf, label = "2012: WA, CO", 
           vjust = 1.5, hjust = -0.1, color = "red", size = 4) +
  annotate("text", x = 2014, y = Inf, label = "2014: AK, OR", 
           vjust = 1.5, hjust = -0.1, color = "orange", size = 4) +
  labs(title = "Marijuana Use Rates: Early Adopters vs Other States (18-25)",
       y = "Rate per 1000",
       color = "State Group") +
  scale_color_manual(values = c("Early Adopters" = "#009E73", "Other States" = "#999999")) +
  theme_minimal(base_size = 14)
```

## Pre vs Post Legalization Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  filter(State %in% early_states) %>%
  mutate(Period = ifelse(Year < 2012, "Pre-2012", 
                         ifelse(Year >= 2012 & Year < 2014, "2012-2013",
                                "Post-2014")),
         Period = factor(Period, levels = c("Pre-2012", "2012-2013", "Post-2014"))) %>%
  ggplot(aes(x = Year, y = `Rates.Marijuana.Used Past Month.18-25`, color = State)) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = 2012, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "orange", size = 1) +
  facet_wrap(~State, ncol = 2) +
  labs(title = "Marijuana Use in Early Adopter States (18-25)",
       y = "Rate per 1000") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Rate of Change Analysis

```{r fig.width=12, fig.height=7}
change_data <- drugs %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(State, Group) %>%
  arrange(Year) %>%
  mutate(Rate_Change = `Rates.Marijuana.Used Past Month.18-25` - 
                       lag(`Rates.Marijuana.Used Past Month.18-25`)) %>%
  ungroup()

change_data %>%
  filter(!is.na(Rate_Change)) %>%
  group_by(Year, Group) %>%
  summarise(Avg_Change = mean(Rate_Change, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = Year, y = Avg_Change, fill = Group)) +
  geom_col(position = "dodge") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_vline(xintercept = 2012, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "orange", size = 1) +
  labs(title = "Year-over-Year Change in Marijuana Rates (18-25)",
       y = "Change in Rate per 1000",
       fill = "State Group") +
  scale_fill_manual(values = c("Early Adopters" = "#009E73", "Other States" = "#999999")) +
  theme_minimal(base_size = 14)
```

## Question 2: Findings

::: {style="font-size: 0.75em;"}
```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

pre_2012 <- drugs %>%
  filter(Year < 2012) %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE))

post_2014 <- drugs %>%
  filter(Year >= 2014) %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE))

cat("Pre-2012 Average Rates (per 1000):\n")
cat("  Early Adopters:", round(pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"], 2), "\n")
cat("  Other States:", round(pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 2), "\n")
cat("  Difference:", round(pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"] - 
                           pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 2), "\n\n")

cat("Post-2014 Average Rates (per 1000):\n")
cat("  Early Adopters:", round(post_2014$Avg_Rate[post_2014$Group == "Early Adopters"], 2), "\n")
cat("  Other States:", round(post_2014$Avg_Rate[post_2014$Group == "Other States"], 2), "\n")
cat("  Difference:", round(post_2014$Avg_Rate[post_2014$Group == "Early Adopters"] - 
                           post_2014$Avg_Rate[post_2014$Group == "Other States"], 2), "\n\n")

increase_early <- post_2014$Avg_Rate[post_2014$Group == "Early Adopters"] - 
                  pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"]
increase_other <- post_2014$Avg_Rate[post_2014$Group == "Other States"] - 
                  pre_2012$Avg_Rate[pre_2012$Group == "Other States"]

cat("Change from Pre-2012 to Post-2014:\n")
cat("  Early Adopters: +", round(increase_early, 2), "(", 
    round(100 * increase_early / pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"], 1), "% increase)\n")
cat("  Other States: +", round(increase_other, 2), "(", 
    round(100 * increase_other / pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 1), "% increase)\n")
```

**Interpretation:**

-   If early adopters show larger % increase → **Causal effect of legalization**
-   If increases are similar → **National trajectory** (legalization followed existing demand)
-   Baseline differences suggest legalization reflected **local preferences**
:::

## Question 2: Conclusion

::: {style="font-size: 0.85em;"}
**The Numbers Tell the Story:**

**Pre-legalization (before 2012):** - Early adopters: 220.6 per 1000 - Other states: 174.7 per 1000 - Already **26% higher** in early adopter states

**Post-legalization (2014+):** - Early adopters: 294.4 per 1000\
- Other states: 211.6 per 1000 - Gap widened to **39% higher**

**Key Finding:** - Early adopters: **33.5% increase** (+73.9 per 1000) - Other states: **21.1% increase** (+36.8 per 1000) - Early adopters increased **58% faster** than other states

**Answer:** Legalization appears to have **accelerated** marijuana usage beyond the national trajectory. Early adopter states not only started with higher baseline usage (suggesting pre-existing demand), but also experienced significantly steeper increases post-legalization.
:::

## Statistical Testing: Did Legalization Matter?

::: {style="font-size: 0.75em;"}
**Hypothesis Test Options:**

**1. Two-Sample t-test on Growth Rates** - Compare % increases: 33.5% vs 21.1% - Tests if growth rates significantly differ

**2. Interaction Effect (Regression)** - Model: Rate \~ Year + Group + Year×Group - Tests if the slope (time trend) differs by group
:::

## Two-Sample t-test on Growth Rates

::: {style="font-size: 0.75em;"}
Let's run a simple t-test:

$$
H_0:\ \mu_{\text{Early growth}} = \mu_{\text{Other growth}}
$$

$$
H_1:\ \mu_{\text{Early growth}} > \mu_{\text{Other growth}}
$$

```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

growth_data <- drugs %>%
  filter(Year %in% c(2010, 2011, 2015, 2016)) %>%
  mutate(Group = ifelse(State %in% early_states, "Early", "Other"),
         Period = ifelse(Year <= 2011, "Pre", "Post")) %>%
  group_by(State, Group, Period) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Period, values_from = Avg_Rate) %>%
  mutate(Growth = Post - Pre)

t_test <- t.test(Growth ~ Group, data = growth_data, alternative = "greater")

cat("Two-Sample t-test: Early Adopters vs Other States\n")
cat("Mean growth Early Adopters:", round(mean(growth_data$Growth[growth_data$Group == "Early"], na.rm = TRUE), 2), "\n")
cat("Mean growth Other States:", round(mean(growth_data$Growth[growth_data$Group == "Other"], na.rm = TRUE), 2), "\n")
cat("t-statistic:", round(t_test$statistic, 3), "\n")
cat("p-value:", round(t_test$p.value, 4), "\n")
cat("\nConclusion:", ifelse(t_test$p.value < 0.05, 
                            "Growth is SIGNIFICANTLY higher in early adopter states (p < 0.05)", 
                            "No significant difference detected"))
```
:::

## Bayesian Approach: Question 2

**Why Bayesian?**

-   Quantify **uncertainty** in the effect of legalization
-   Get **probability distributions** instead of just point estimates
-   Make probabilistic statements: "What's the probability that early adopters increased more?"
-   Incorporate **prior knowledge** about drug use trends

## Proposed Model

::: {style="font-size: 0.7em;"}
$$
Rate_{it} \sim \text{Normal}(\mu_{it}, \sigma)
$$

$$
\mu_{it} = \alpha_i + \beta_1(Year_t - 2012) + \beta_2 Post_{it}
$$
:::

**Priors:**

::: {style="font-size: 0.6em;"}
$$
\alpha_i \sim \text{Normal}(\alpha_0, \tau_\alpha)
$$

$$
\alpha_0 \sim \text{Normal}(180, 50)
$$

$$
\tau_\alpha \sim \text{Half-Normal}(30)
$$

$$
\beta_1 \sim \text{Normal}(0, 10)
$$

$$
\beta_2 \sim \text{Normal}(0, 20)
$$

$$
\sigma \sim \text{Half-Normal}(25)
$$
:::

## Model Interpretation

::: {style="font-size: 0.6em;"}
**Key Parameter:** $β_2$ (Post-Legalization Effect)

If $β_2 > 0$, marijuana-use rates increased after legalization in treated states beyond the broader national trend. Posterior $P(β_2 > 0 \mid data)$ gives the probability that this effect is positive under the model.

**Model Notes:**

-   $Post_{it}$: state-specific post-legalization indicator, since legalization occurred in different years across states

-   $α_i$: state-specific baseline marijuana-use level

-   Time is centered at 2012 for easier interpretation of the intercept and trend

**Posterior Predictive Checks:** 

-   Simulate replicated state-year rates from the posterior

-   Compare simulated values to observed trends over time 

-   Check whether the model captures national trends, state-level variation, and post-legalization shifts
:::

```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

drugs_q2 <- drugs %>%
  mutate(
    Rate = `Rates.Marijuana.Used Past Month.18-25`,
    EarlyAdopter = if_else(State %in% early_states, 1, 0),
    LegalYear = case_when(
      State == "Washington" ~ 2012,
      State == "Colorado" ~ 2012,
      State == "Alaska" ~ 2014,
      State == "Oregon" ~ 2014,
      TRUE ~ 9999
    ),
    Post = if_else(Year >= LegalYear, 1, 0),
    Year_c = Year - 2012
  ) %>%
  filter(!is.na(Rate)) %>%
  select(State, Year, Rate, EarlyAdopter, Post, Year_c)
```

```{r}
if (file.exists("legalization_model.rds")) {
  legalization_model <- readRDS("legalization_model.rds")
} else {
  legalization_model <- stan_glmer(
    Rate ~ Year_c + Post + (1 | State),
    data = drugs_q2,
    family = gaussian(),
    prior_intercept = normal(180, 50, autoscale = FALSE),
    prior = normal(0, 20, autoscale = FALSE),
    prior_aux = exponential(1, autoscale = TRUE),
    chains = 4,
    iter = 5000*2,
    seed = 84735,
    refresh = 0
  )
  saveRDS(legalization_model, "legalization_model.rds")
}


```

## MCMC Trace
```{r}
mcmc_trace(
  as.array(legalization_model),
  pars = c("(Intercept)", "Year_c", "Post", "sigma", "Sigma[State:(Intercept),(Intercept)]"),
  size = 0.2
)
```

## MCMC Density Overlay

```{r}
mcmc_dens_overlay(
  as.array(legalization_model),
  pars = c("(Intercept)", "Year_c", "Post", "sigma")
)
```

## Bayesian Analysis Results

::: {style="font-size: 0.75em;"}
**What we'd report (with actual Bayesian analysis):**

```         
Posterior Summary for β₄ (Interaction Effect):
  Mean: 5.2 per 1000 per year
  95% Credible Interval: [2.1, 8.7]
  P(β₄ > 0 | data) = 0.997
  
Interpretation: 
- 99.7% probability that legalization caused ADDITIONAL increase in early adopter states
- Expected extra increase: ~5.2 per 1000 per year beyond national trend
- We are 95% confident the effect is between 2.1 and 8.7
```

**Posterior Predictive Check:** - Generate 1000 datasets from posterior - Compare simulated trends to actual data - Check: Do simulated rates match observed variance? State patterns? Time trends?

**Advantages over Frequentist:** - Direct probability statements about effects - Uncertainty quantification for predictions - Can incorporate domain knowledge via priors - Natural handling of hierarchical structure (states within groups)
:::

## Implementing the Bayesian Model

::: {style="font-size: 0.65em;"}
**Simple Bayesian Analysis (Using Conjugate Priors):**

```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

pre_data <- drugs %>%
  filter(Year < 2012) %>%
  mutate(Group = ifelse(State %in% early_states, "Early", "Other"))

post_data <- drugs %>%
  filter(Year >= 2014) %>%
  mutate(Group = ifelse(State %in% early_states, "Early", "Other"))

early_pre <- pre_data %>% filter(Group == "Early") %>% pull(`Rates.Marijuana.Used Past Month.18-25`)
early_post <- post_data %>% filter(Group == "Early") %>% pull(`Rates.Marijuana.Used Past Month.18-25`)
other_pre <- pre_data %>% filter(Group == "Other") %>% pull(`Rates.Marijuana.Used Past Month.18-25`)
other_post <- post_data %>% filter(Group == "Other") %>% pull(`Rates.Marijuana.Used Past Month.18-25`)

change_early <- mean(early_post, na.rm = TRUE) - mean(early_pre, na.rm = TRUE)
change_other <- mean(other_post, na.rm = TRUE) - mean(other_pre, na.rm = TRUE)
diff_in_diff <- change_early - change_other

set.seed(123)
n_sims <- 10000

sim_change_early <- rnorm(n_sims, mean = change_early, sd = sd(early_post - mean(early_pre), na.rm = TRUE) / sqrt(length(early_post)))
sim_change_other <- rnorm(n_sims, mean = change_other, sd = sd(other_post - mean(other_pre), na.rm = TRUE) / sqrt(length(other_post)))
posterior_diff <- sim_change_early - sim_change_other

cat("Bayesian Difference-in-Differences Analysis\n")
cat("==========================================\n\n")
cat("Posterior Mean (Extra effect for early adopters):", round(mean(posterior_diff), 2), "\n")
cat("95% Credible Interval:", round(quantile(posterior_diff, 0.025), 2), "to", round(quantile(posterior_diff, 0.975), 2), "\n")
cat("Posterior P(Effect > 0 | data):", round(mean(posterior_diff > 0), 4), "\n")
cat("Posterior P(Effect > 20 | data):", round(mean(posterior_diff > 20), 4), "\n\n")
cat("Interpretation: There is a", round(mean(posterior_diff > 0) * 100, 1), "% probability that\n")
cat("legalization caused additional increases in early adopter states.\n")
```

```{r fig.width=8, fig.height=3.5}
hist(posterior_diff, breaks = 50, col = "skyblue", border = "white",
     main = "Posterior Distribution: Extra Effect of Legalization",
     xlab = "Additional Change in Early Adopters (per 1000)",
     freq = FALSE)
abline(v = mean(posterior_diff), col = "red", lwd = 2, lty = 2)
abline(v = quantile(posterior_diff, c(0.025, 0.975)), col = "darkred", lwd = 2, lty = 3)
abline(v = 0, col = "black", lwd = 2)
legend("topright", c("Posterior Mean", "95% Credible Interval", "No Effect"),
       col = c("red", "darkred", "black"), lty = c(2, 3, 1), lwd = 2)
```
:::

## Comparing the Approaches

::: {style="font-size: 0.75em;"}
**Descriptive Analysis (Earlier):** - Early adopters: +73.9 per 1000 (33.5% increase) - Other states: +36.8 per 1000 (21.1% increase)\
- **Difference: \~37 per 1000**

**Bayesian Analysis (Just now):** - **Posterior mean: 37.05** per 1000 extra effect - 95% Credible Interval: \[19.23, 54.5\] - P(Effect \> 0) = 100%

**Why are they similar?**

They're measuring the **same thing** (difference-in-differences), but the Bayesian approach adds:

1.  **Uncertainty quantification**: Not just "37", but "37 ± uncertainty"
2.  **Probabilistic statements**: "100% probability of positive effect" vs "p \< 0.05"
3.  **Credible intervals**: "95% chance true value is in \[19, 55\]" vs "we reject null hypothesis"
4.  **Full distribution**: See the entire range of plausible values (the histogram)

**Key insight:** The Bayesian analysis **confirms** our descriptive finding but gives us **stronger evidence** with quantified uncertainty. The posterior mean (37.05) matches our difference-in-differences (37) because both are based on the same data, but Bayesian tells us how confident we should be.
:::
