---
title: "Best Presentation"
subtitle: "Stats 115 Final Project"
author: "Charles Gao<br>Andrew Hwang<br>Jason Huffstutter<br>Andrew Hahm"
format: revealjs
editor: visual
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)

drugs <- read_csv("data/drugs.csv")
```

## Background & Motivation
TO DO

## Our Research Question
-  How do years of economic hardship (2008) affect alcohol rates across all age groups?
-  Did early state-level marijuana legalization in 2012 (Washington, Colorado) and 2014 (Alaska, Oregon) represent a causal driver of increased usage, or did usage rates simply follow the national trajectory?

## Dataset Overview

::: {style="font-size: 0.7em;"}
```{r}
cat("Dimensions:", nrow(drugs), "observations ×", ncol(drugs), "variables\n")
cat("Years:", min(drugs$Year), "-", max(drugs$Year), "\n")
cat("States:", length(unique(drugs$State)), "\n")
```

**Key Variables:**

-   **Demographics**: State, Year, Population by age group (12-17, 18-25, 26+)
-   **Substances**: Alcohol, Marijuana, Tobacco, Cocaine
-   **Metrics**: Totals (counts) and Rates (per 1000 population)
-   **Time frames**: Past month, past year usage

**Age Groups:**

-   12-17 (adolescents)
-   18-25 (young adults)
-   26+ (adults)
:::

## Marijuana Trends by Age

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Marijuana.Used Past Month.12-17`,
         `Rates.Marijuana.Used Past Month.18-25`,
         `Rates.Marijuana.Used Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Marijuana Use Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group shows an increasing rate per 1,000 over time and remains significantly higher than the other two age groups.
:::

## Alcohol Disorder Trends

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Alcohol.Use Disorder Past Year.12-17`,
         `Rates.Alcohol.Use Disorder Past Year.18-25`,
         `Rates.Alcohol.Use Disorder Past Year.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Alcohol Use Disorder Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group still shows the highest alcohol disorder rate per 1,000, but rates for all three age groups gradually decrease over time.
:::

## Tobacco Trends

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Tobacco.Cigarette Past Month.12-17`,
         `Rates.Tobacco.Cigarette Past Month.18-25`,
         `Rates.Tobacco.Cigarette Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Tobacco Use Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group shows the highest tobacco usage rate per 1,000. Rates for all three age groups gradually decline over time, but at different speeds, with the 18–25 group eventually converging with the 26+ group.
:::

## Top States: Marijuana Use

```{r fig.width=10, fig.height=7}
drugs %>%
  filter(Year == max(Year)) %>%
  arrange(desc(`Rates.Marijuana.Used Past Month.18-25`)) %>%
  slice(1:15) %>%
  ggplot(aes(x = reorder(State, `Rates.Marijuana.Used Past Month.18-25`), 
             y = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 15 States: Marijuana Use (18-25)", x = "", y = "Rate per 1000")
```

## Distribution: Marijuana Use

```{r fig.width=10, fig.height=6}
drugs %>%
  ggplot(aes(x = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_histogram(bins = 30, fill = "seagreen", alpha = 0.7) +
  geom_density(aes(y = after_stat(count)), size = 1.2, color = "darkgreen") +
  labs(title = "Distribution of Marijuana Use Rates (18-25)", x = "Rate per 1000")
```

## Correlation Matrix

```{r fig.width=12, fig.height=10}
cor_data <- drugs %>%
  select(Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Alcohol = `Rates.Alcohol.Use Disorder Past Year.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`)

cor_matrix <- cor(cor_data, use = "complete.obs")

cor_df <- as.data.frame(as.table(cor_matrix))
names(cor_df) <- c("Var1", "Var2", "Correlation")

ggplot(cor_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white", size = 1) +
  geom_text(aes(label = round(Correlation, 2)), size = 8) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Between Substances (18-25)") +
  theme_minimal(base_size = 16) +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.text.y = element_text(size = 14))
```

::: {style="font-size: 0.8em;"}
**Two distinct camps:** Cocaine & Marijuana users (r = 0.55) vs Alcohol & Tobacco users (r = 0.73)

**Key insight:** Marijuana and Tobacco negatively correlated (r = -0.27) - both are smoking activities, individuals choose one over the other
:::

## All Substance Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  select(Year, 
         Alcohol = `Rates.Alcohol.Use Past Month.18-25`,
         Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`) %>%
  pivot_longer(cols = -Year, names_to = "Substance", values_to = "Rate") %>%
  ggplot(aes(x = Year, y = Rate, color = Substance)) +
  geom_smooth(se = TRUE) +
  facet_wrap(~Substance, scales = "free_y") +
  labs(title = "All Substances (18-25 Age Group)") +
  theme(legend.position = "none")
```
