---
title: "Presentation"
format: revealjs
editor: visual
---

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)

drugs <- read_csv("data/drugs.csv")
```

## Dataset Overview

```{r}
dim(drugs)
head(drugs, 3)
```

## Marijuana Trends by Age

```{r fig.width=10, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Marijuana.Used Past Month.12-17`,
         `Rates.Marijuana.Used Past Month.18-25`,
         `Rates.Marijuana.Used Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Marijuana Use Over Time", y = "Rate per 1000")
```

## Alcohol Disorder Trends

```{r fig.width=10, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Alcohol.Use Disorder Past Year.12-17`,
         `Rates.Alcohol.Use Disorder Past Year.18-25`,
         `Rates.Alcohol.Use Disorder Past Year.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Alcohol Use Disorder Over Time", y = "Rate per 1000")
```

## Tobacco Trends

```{r fig.width=10, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Tobacco.Cigarette Past Month.12-17`,
         `Rates.Tobacco.Cigarette Past Month.18-25`,
         `Rates.Tobacco.Cigarette Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Tobacco Use Over Time", y = "Rate per 1000")
```

## Top States: Marijuana Use

```{r fig.width=10, fig.height=7}
drugs %>%
  filter(Year == max(Year)) %>%
  arrange(desc(`Rates.Marijuana.Used Past Month.18-25`)) %>%
  slice(1:15) %>%
  ggplot(aes(x = reorder(State, `Rates.Marijuana.Used Past Month.18-25`), 
             y = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 15 States: Marijuana Use (18-25)", x = "", y = "Rate per 1000")
```

## Distribution: Marijuana Use

```{r fig.width=10, fig.height=6}
drugs %>%
  ggplot(aes(x = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_histogram(bins = 30, fill = "seagreen", alpha = 0.7) +
  geom_density(aes(y = after_stat(count)), size = 1.2, color = "darkgreen") +
  labs(title = "Distribution of Marijuana Use Rates (18-25)", x = "Rate per 1000")
```

## Correlation Matrix

```{r fig.width=10, fig.height=8}
cor_data <- drugs %>%
  select(Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Alcohol = `Rates.Alcohol.Use Disorder Past Year.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`)

cor_matrix <- cor(cor_data, use = "complete.obs")

cor_df <- as.data.frame(as.table(cor_matrix))
names(cor_df) <- c("Var1", "Var2", "Correlation")

ggplot(cor_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white") +
  geom_text(aes(label = round(Correlation, 2)), size = 5) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Between Substances (18-25)") +
  theme_minimal() +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## All Substance Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  select(Year, 
         Alcohol = `Rates.Alcohol.Use Past Month.18-25`,
         Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`) %>%
  pivot_longer(cols = -Year, names_to = "Substance", values_to = "Rate") %>%
  ggplot(aes(x = Year, y = Rate, color = Substance)) +
  geom_smooth(se = TRUE) +
  facet_wrap(~Substance, scales = "free_y") +
  labs(title = "All Substances (18-25 Age Group)") +
  theme(legend.position = "none")
```

## Initial Analysis (1)

From Marijuana Trends by age, Alcohol Disorder Trends, Tobacco Trends (The complete pool of all states)

-   Age group from 18-25 display increasing rate per 1000 over time, and significantly higher compare to the other two age groups.

-   Age group from 18-25 still present higher disorder rate per 1000, but the rate for all 3 age groups are gradually decreasing over time.

-   Age group from 18-25 display higher higher tobacco usage rate per 1000. The rate for all 3 age groups are gradually decreasing but at different speed, where age group 18-25 meets the usage rate with age group from 26+
