---
title: "Legalization<br>& Cannabis Use"
subtitle: "Stats 115 Final Project"
author: "Charles Gao<br>Andrew Hwang<br>Jason Huffstutter<br>Andrew Hahm"
format: 
  revealjs: 
    self-contained: true
    embed-resources: true
editor: visual
---

```{r}
suppressWarnings(suppressMessages({
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(stringr)
library(bayesrules)
library(rstanarm)
library(bayesplot)
library(tidyverse)
library(broom.mixed)
library(tidybayes)
}))
drugs <- read_csv("../data/drugs.csv")
```

## Our Research Question

-   Was early state-level marijuana legalization in 2012 (Washington, Colorado) and 2014 (Alaska, Oregon) associated with higher marijuana-use rates among ages 18–25, beyond the broader national trend?

## Background & Motivation

::: {style="font-size: 0.75em;"}
Why should we care about this topic?

-   Marijuana legalization is a policy issue in the United States that goes beyond red/blue politics, as it reflects family dynamics, community experiences, and personal values.

-   A key concern is whether legalization is associated with significantly higher marijuana use, especially among young adults.

Who will these results help in the future?

-   Policymakers can use these results to better understand how marijuana regulation can affect younger demographics.

-   Public health groups and schools can use this information to better educate and improve prevention efforts.

-   Individuals can gain knowledge to create more informed opinions.
:::

## Dataset Overview

::: {style="font-size: 0.7em;"}
```{r}
cat("Dimensions:", nrow(drugs), "observations ×", ncol(drugs), "variables\n")
cat("Years:", min(drugs$Year), "-", max(drugs$Year), "\n")
cat("States:", length(unique(drugs$State)), "(includes Washington D.C.)\n")
```

**Key Variables:**

-   **Demographics**: State, Year, Population by age group
-   **Substances**: Marijuana, Tobacco, Alcohol, Cocaine
-   **Metrics**: Totals (counts) and Rates (per 1000 people)
-   **Time frames**: Past month, past year usage

**Age Groups:**

-   12-17 (adolescents)
-   18-25 (young adults)
-   26+ (adults)
:::

## Marijuana Trends by Age

```{r fig.width=12, fig.height=6}
drugs %>%
  select(Year, 
         `Rates.Marijuana.Used Past Month.12-17`,
         `Rates.Marijuana.Used Past Month.18-25`,
         `Rates.Marijuana.Used Past Month.26+`) %>%
  pivot_longer(cols = -Year, names_to = "Age_Group", values_to = "Rate") %>%
  mutate(Age_Group = str_extract(Age_Group, "12-17|18-25|26\\+")) %>%
  ggplot(aes(x = Year, y = Rate, color = Age_Group)) +
  geom_smooth(se = TRUE) +
  labs(title = "Marijuana Use Over Time", y = "Rate per 1000")
```

::: {style="font-size: 60%;"}
The 18–25 age group shows an increasing rate over time and remains significantly higher than the other two age groups.
:::

## States Ranked by Marijuana Use (2018)

```{r fig.width=10, fig.height=5}
drugs %>%
  filter(Year == max(Year)) %>%
  arrange(desc(`Rates.Marijuana.Used Past Month.18-25`)) %>%
  slice(1:15) %>%
  ggplot(aes(x = reorder(State, `Rates.Marijuana.Used Past Month.18-25`), 
             y = `Rates.Marijuana.Used Past Month.18-25`)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(title = "Top 15 States: Marijuana Use (18-25)", x = "", y = "Rate per 1000")
```

## Correlation Matrix

```{r fig.width=12, fig.height=10}
cor_data <- drugs %>%
  select(Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Alcohol = `Rates.Alcohol.Use Disorder Past Year.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`)

cor_matrix <- cor(cor_data, use = "complete.obs")

cor_df <- as.data.frame(as.table(cor_matrix))
names(cor_df) <- c("Var1", "Var2", "Correlation")

ggplot(cor_df, aes(x = Var1, y = Var2, fill = Correlation)) +
  geom_tile(color = "white", size = 1) +
  geom_text(aes(label = round(Correlation, 2)), size = 8) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", 
                       midpoint = 0, limits = c(-1, 1)) +
  labs(title = "Correlation Between Substances (18-25)") +
  theme_minimal(base_size = 16) +
  theme(axis.title = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 14),
        axis.text.y = element_text(size = 14))
```

::: {style="font-size: 0.75em;"}
**Two distinct camps:** Cocaine & Marijuana users (r = 0.55) vs Alcohol & Tobacco users (r = 0.73)

**Key insight:** Marijuana and Tobacco negatively correlated (r = -0.27) - both are smoking activities, individuals choose one over the other
:::

## All Substance Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  select(Year, 
         Alcohol = `Rates.Alcohol.Use Past Month.18-25`,
         Marijuana = `Rates.Marijuana.Used Past Month.18-25`,
         Tobacco = `Rates.Tobacco.Cigarette Past Month.18-25`,
         Cocaine = `Rates.Illicit Drugs.Cocaine Used Past Year.18-25`) %>%
  pivot_longer(cols = -Year, names_to = "Substance", values_to = "Rate") %>%
  ggplot(aes(x = Year, y = Rate, color = Substance)) +
  geom_smooth(se = TRUE) +
  facet_wrap(~Substance, scales = "free_y") +
  labs(title = "All Substances (18-25 Age Group)") +
  theme(legend.position = "none")
```

## Revisiting our Question

::: {style="font-size: 0.8em;"}
-   Was early state-level marijuana legalization in 2012 (Washington, Colorado) and 2014 (Alaska, Oregon) associated with higher marijuana-use rates among ages 18–25, beyond the broader national trend?

Our exploratory analysis suggested that marijuana use among **18–25-year-olds** rose in the years following legalization. We then asked whether early-adopter states experienced an additional post-legalization increase relative to the broader national pattern, or whether usage was simply rising along with the national trajectory.
:::

## Legalization Timeline

::: {style="font-size: 0.75em;"}
**Early Adopter States:**

-   **2012**: Washington & Colorado legalized recreational marijuana
-   **2014**: Alaska & Oregon legalized recreational marijuana

**Our Analysis:**

Compare marijuana usage rates in these 4 early-adopter states vs. all other states to identify:

1.  Pre-legalization baseline differences
2.  Post-legalization divergence (if any)
3.  Whether early-adopter states showed larger post-legalization increases than other states
:::

## Marijuana Rates: Early Adopters vs Others

```{r fig.width=12, fig.height=5}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

drugs %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Year, Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = Year, y = Avg_Rate, color = Group, group = Group)) +
  geom_line(size = 1.5) +
  geom_point(size = 3) +
  geom_vline(xintercept = 2012, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "orange", size = 1) +
  annotate("text", x = 2012, y = Inf, label = "2012: WA, CO", 
           vjust = 1.5, hjust = -0.1, color = "red", size = 4) +
  annotate("text", x = 2014, y = Inf, label = "2014: AK, OR", 
           vjust = 1.5, hjust = -0.1, color = "orange", size = 4) +
  labs(title = "Marijuana Use Rates: Early Adopters vs Other States (18-25)",
       y = "Rate per 1000",
       color = "State Group") +
  scale_color_manual(values = c("Early Adopters" = "#009E73", "Other States" = "#999999")) +
  theme_minimal(base_size = 14)
```

## Pre vs Post Legalization Trends

```{r fig.width=12, fig.height=7}
drugs %>%
  filter(State %in% early_states) %>%
  mutate(Period = ifelse(Year < 2012, "Pre-2012", 
                         ifelse(Year >= 2012 & Year < 2014, "2012-2013",
                                "Post-2014")),
         Period = factor(Period, levels = c("Pre-2012", "2012-2013", "Post-2014"))) %>%
  ggplot(aes(x = Year, y = `Rates.Marijuana.Used Past Month.18-25`, color = State)) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5) +
  geom_vline(xintercept = 2012, linetype = "dashed", color = "red", size = 1) +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "orange", size = 1) +
  facet_wrap(~State, ncol = 2) +
  labs(title = "Marijuana Use in Early Adopter States (18-25)",
       y = "Rate per 1000") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "none")
```

## Findings

::: {style="font-size: 0.6em;"}
```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

pre_2012 <- drugs %>%
  filter(Year < 2012) %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE))

post_2014 <- drugs %>%
  filter(Year >= 2014) %>%
  mutate(Group = ifelse(State %in% early_states, "Early Adopters", "Other States")) %>%
  group_by(Group) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE))

cat("Pre-2012 Average Rates (per 1000):\n")
cat("  Early Adopters:", round(pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"], 2), "\n")
cat("  Other States:", round(pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 2), "\n")
cat("  Difference:", round(pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"] - 
                           pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 2), "\n\n")

cat("Post-2014 Average Rates (per 1000):\n")
cat("  Early Adopters:", round(post_2014$Avg_Rate[post_2014$Group == "Early Adopters"], 2), "\n")
cat("  Other States:", round(post_2014$Avg_Rate[post_2014$Group == "Other States"], 2), "\n")
cat("  Difference:", round(post_2014$Avg_Rate[post_2014$Group == "Early Adopters"] - 
                           post_2014$Avg_Rate[post_2014$Group == "Other States"], 2), "\n\n")

increase_early <- post_2014$Avg_Rate[post_2014$Group == "Early Adopters"] - 
                  pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"]
increase_other <- post_2014$Avg_Rate[post_2014$Group == "Other States"] - 
                  pre_2012$Avg_Rate[pre_2012$Group == "Other States"]

cat("Change from Pre-2012 to Post-2014:\n")
cat("  Early Adopters: +", round(increase_early, 2), "(", 
    round(100 * increase_early / pre_2012$Avg_Rate[pre_2012$Group == "Early Adopters"], 1), "% increase)\n")
cat("  Other States: +", round(increase_other, 2), "(", 
    round(100 * increase_other / pre_2012$Avg_Rate[pre_2012$Group == "Other States"], 1), "% increase)\n")
```

*Note: This analysis uses all pre-2012 vs all post-2014 data for comprehensive coverage.*

**Interpretation:**

-   Larger increases in early-adopter states are consistent with a positive post-legalization association, but do **not establish causality by themselves**
-   Similar increases across groups would be more consistent with a broader national trend
-   Baseline differences suggest early-adopter states may already have had stronger underlying demand or different pre-existing preferences
:::

## Statistical Testing: Did Legalization Matter?

::: {style="font-size: 0.7em;"}
**Modeling Strategy:**

Because legalization occurred at different times across the treated states, we define a **state-specific post-legalization indicator**:

-   $Post_{it}=1$ if state $i$ is in a legalized period in year $t$

-   $Post_{it}=0$ otherwise

This means $Post_{it}$ already represents the treatment exposure for early-adopter states after legalization. It is effectively a **treated × post** indicator, so we do **not** need a separate interaction term such as $EarlyAdopter_i \times Post_t$ in this model.
:::

## Two-Sample t-test on Growth Rates

::: {style="font-size: 0.75em;"}
Let's first run a simple t-test:

$$
H_0:\ \mu_{\text{Early growth}} = \mu_{\text{Other growth}}
$$

$$
H_1:\ \mu_{\text{Early growth}} > \mu_{\text{Other growth}}
$$

```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

growth_data <- drugs %>%
  filter(Year %in% c(2010, 2011, 2015, 2016)) %>%
  mutate(Group = ifelse(State %in% early_states, "Early", "Other"),
         Period = ifelse(Year <= 2011, "Pre", "Post")) %>%
  group_by(State, Group, Period) %>%
  summarise(Avg_Rate = mean(`Rates.Marijuana.Used Past Month.18-25`, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = Period, values_from = Avg_Rate) %>%
  mutate(Growth = Post - Pre)

t_test <- t.test(Growth ~ Group, data = growth_data, alternative = "greater")

cat("Two-Sample t-test: Early Adopters vs Other States\n")
cat("Mean growth Early Adopters:", round(mean(growth_data$Growth[growth_data$Group == "Early"], na.rm = TRUE), 2), "\n")
cat("Mean growth Other States:", round(mean(growth_data$Growth[growth_data$Group == "Other"], na.rm = TRUE), 2), "\n")
cat("t-statistic:", round(t_test$statistic, 3), "\n")
cat("p-value:", round(t_test$p.value, 4), "\n")
cat("\nConclusion:", ifelse(t_test$p.value < 0.05, 
                            "Average growth was statistically higher in early-adopter states", 
                            "No statistically significant difference in average growth was detected"))
```

**Note:** This t-test uses 2010-2011 (pre) vs 2015-2016 (post) for a clean comparison around the legalization dates.
:::

## Proposed Model

::: {style="font-size: 0.7em;"}
$$
Rate_{it} \sim \text{Normal}(\mu_{it}, \sigma)
$$

$$
\mu_{it} = \alpha_i + \lambda_t + \beta Post_{it}
$$
:::

**Where:**

::: {style="font-size: 0.65em;"}
-   $Rate_{it}$ is the marijuana-use rate for state $i$ in year $t$
-   $\alpha_i$ is the state-specific baseline level
-   $\lambda_t$ is the year-specific effect shared by all states
-   $Post_{it}$ is a state-specific post-legalization indicator
:::

**Why use year effects instead of a linear year trend?**

::: {style="font-size: 0.65em;"}
A single linear time term assumes marijuana-use rates changed at a constant national rate over time. Using year effects is more flexible: it allows each year to have its own national shift, which better captures non-linear changes in marijuana use over the sample period.
:::

## Model Interpretation

::: {style="font-size: 0.6em;"}
**Key Parameter:** $β$ ($Post_{it}$)

If $β > 0$, then marijuana-use rates were higher in treated states during their post-legalization years, after accounting for:

-   baseline differences across states through $α_i$
-   nationwide year-to-year shocks through $λ_t$

Because $Post_{it}$ is defined separately for each state based on its own legalization year, it already serves as the treatment indicator for “treated state after legalization.”

**Parameter meanings:**

-   $α_i$: state-specific baseline marijuana-use level
-   $λ_t$: year-specific national effect
-   $β$: additional post-legalization shift in treated states
:::

```{r}
early_states <- c("Washington", "Colorado", "Alaska", "Oregon")

drugs_q2 <- drugs %>%
  mutate(
    Rate = `Rates.Marijuana.Used Past Month.18-25`,
    LegalYear = case_when(
      State == "Washington" ~ 2012,
      State == "Colorado" ~ 2012,
      State == "Alaska" ~ 2014,
      State == "Oregon" ~ 2014,
      TRUE ~ 9999
    ),
    Post = if_else(Year >= LegalYear, 1, 0)
  ) %>%
  filter(!is.na(Rate)) %>%
  select(State, Year, Rate, Post)
```

```{r}
if (file.exists("legalization_model.rds")) {
  legalization_model <- readRDS("legalization_model.rds")
} else {
  legalization_model <- stan_glmer(
    Rate ~ factor(Year) + Post + (1 | State),
    data = drugs_q2,
    family = gaussian(),
    prior_intercept = normal(180, 50, autoscale = FALSE),
    prior = normal(0, 20, autoscale = FALSE),
    prior_aux = exponential(1, autoscale = TRUE),
    chains = 4,
    iter = 5000 * 2,
    seed = 84735,
    refresh = 0
  )
  saveRDS(legalization_model, "legalization_model.rds")
}


```

## MCMC Trace

```{r}
mcmc_trace(
  as.array(legalization_model),
  pars = c("(Intercept)", "Post", "sigma", "Sigma[State:(Intercept),(Intercept)]"),
  size = 0.2
)
```

::: {style="font-size: 0.6em;"}
The traceplots show good mixing across chains
:::

## MCMC Density Overlay

```{r}
mcmc_dens_overlay(
  as.array(legalization_model),
  pars = c("(Intercept)", "Post", "sigma")
)
```

::: {style="font-size: 0.6em;"}
Density overlays are quite consistent across chains. Intercept chains showed slight variation, but overall the posterior simulation converged well and the MCMC samples are reliable for inference.
:::

## Posterior Predictive Check

```{r}
pp_check(legalization_model)
```

::: {style="font-size: 0.6em;"}
Our simulated data from the posterior match the observed distribution and trend quite well. The model captures the main structure of the marijuana-use data reasonably well.
:::

## Bayesian Analysis Results

::: {style="font-size: 0.7em;"}
**Posterior Summary:**

-   $95\%$ credible interval for $β$ ($Post_{it}$): $[29.02,\ 46.99]$

-   $P(β > 0 \mid data) = 1.00$

**Interpretation:** After accounting for state-level baseline differences and year-specific national effects, the estimated post-legalization coefficient remains clearly positive. Under this model, post-legalization years in treated states were associated with an additional increase of about 29 to 47 rate points in marijuana-use rates among ages 18–25. This estimate reflects a post-legalization level shift rather than a different slope over time.
:::

## Comparing the Approaches

::: {style="font-size: 0.6em;"}
**Descriptive Analysis:**

-   Early-adopter states showed larger increases in marijuana-use rates than other states
-   However, these states also had higher baseline usage before legalization

**Frequentist Test:**

-   The two-sample t-test found higher average growth in early-adopter states
-   However, the difference was not statistically significant

**Bayesian Model:**

-   The model controls for state-level baseline differences and year-specific national effects
-   The post-legalization indicator $Post_{it}$ remained strongly positive: $95\%$ credible interval $[29.02,\ 46.99]$
:::

## Where could we go from here?

::: {style="font-size: 0.7em;"}
-   We could study other factors like dispensary access, advertising, and explore different demographics outside of just age.

-   We could increase the timeframe we are looking at to study long term effects, and potentially combine our findings with other datasets that include information on post 2018 legalization.

-   We could collect more detailed data, such as county-level, instead of only state-level averages.

-   We could explore the link between marijuana and other substances, studying more how their rates are related.
:::

------------------------------------------------------------------------

::: {style="display: flex; justify-content: center; align-items: center; height: 100%; font-size: 3em;"}
**Thank you**
:::
